A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-03-09 sections/model_overview/main.tex
Restructure game theoretic model section

\section{Overview of game theoretic model} \label{sec:model_overview}

@@

\begin{itemize}
    \item \( \lambda_2 \): The rate of non-emergency patients that the EMS 
    receives and distributes to EDs
    \item \( \lambda_{1_i} \): The arrival rate of other patients to ED 
    \(i\in\{A, B\}\)
    \item \( \mu_i \): The service rate of patients at ED \(i\in\{A, B\}\)
    \item \( C_i \): The number of available resources (beds) in ED 
    \(i\in\{A, B\}\)  
    \item \( T_i \): The action that ED \(i\in\{A, B\} \) chooses to play which 
    corresponds to the threshold at which they do not accept EMS patients.
    \item \( N_i \): The total patient capacity of ED \(i\in\{A, B\} \)
    \item \( M_i \): The total parking capacity of ED \(i\in\{A, B\} \)
    \item \( R \): The time target for both EDs
    \item \( \alpha \in [0, 1] \) : \textit{Weighting} of blocking time and 
    lost individuals (equation \ref{eq:obj_distributor})
\end{itemize}

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in Section \ref{sec:model_overview}, the interaction between the two EDs 
is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). This corresponds to reacting to experienced 
delays.

The queueing systems of the hospitals are designed in such a way where they can
accept two types
of individuals. 
Type 2 individuals are defined as non-urgent patients arriving in ambulances 
that are allocated by the ambulance service and 
type 1 individuals are individuals arriving from other sources 
(e.g walk-in patients, urgent patients from either walk-ins or ambulances).
Each hospital may then choose to block type 2 individuals 
when the hospital reaches a certain capacity. 
The strategy sets for each hospital is the set 
\( \{T \in \mathbb{N} \;|\; 1 \leq T \leq N\} \) where \(N \in\{N_A, N_B\}\) are 
the total capacities of hospitals \(A\) and \(B\). We denote the chosen actions 
from the strategy set as \(T_A, T_B\) and call these \textit{threshold}s.

Both hospitals follow a queueing model with two waiting zones for 
individuals. 
The first waiting zone (i.e. the waiting space of the hospital) is where the 
patients queue right before receiving 
their service and has a queue capacity of \( N - C \), where \(N\) is the total 
capacity of the hospital and \(C\) is the number of healthcare 
professionals able to see them.
% TODO: Should this be healthcare professionals or available beds?
The second waiting zone (i.e. the parking space for ambulances) is where 
ambulances, that are sent from the
ambulance service, stay until their patients are allowed to enter the hospital.
The parking space has a capacity of \(M\) and no servers.

@@

\begin{equation}\label{eq:obj_queueing_systems}
    \arg \max_{T_i} \quad -\left( \hat{P} - P(W_i < t) \right)^2
\end{equation}

The third player, the ambulance service, has their own choices to make and 
their own 
goals to satisfy.
The strategy set of the third player is the proportion \(0 \leq p_A \leq 1\) of 
individuals to send to hospital A.
Similarly the proportion of individuals to send to hospital B is given by
\(p_B = 1 - p_A\).
In addition, the ambulance service aims to minimise any potential blockages
that may occur, given the pair of thresholds chosen by the two hospitals.
Thus, its objective is to minimise the blocked time of the individuals 
(\(B_A\) and \(B_B\))
that they send to hospitals \(A\) and \(B\).
Apart from the time being blocked, an additional aspect that may affect the 
decision of the distributor is the proportion of lost individuals 
\(L_A\) and \(L_B\).
Equation \ref{eq:obj_distributor} can be used to capture a mixture 
between the two objectives \(L_i\) and \(B_i\) where \(i \in \{A, B\}\):
\begin{equation*}\label{eq:obj_distributor}
    (p_A, p_B) \quad s.t. \quad 
    \alpha L_A(p_A) + (1 - \alpha) B_A(p_A) = 
    \alpha L_B(p_B) + (1 - \alpha) B_B(p_B)
\end{equation*}

@@

The first 
queueing system \(H_A\) decides on a threshold, then the second system \(H_B\)
chooses its own threshold, without knowing the strategy of \(H_A\), and finally
the distributor makes its choice. Note here that the dotted line represents the
fact that \(H_B\) is unaware of the state of the game when making its own 
decisions.

@@

\begin{equation} \label{eq:payoff_entry}
    U_{T_A, T_B}^i = -\left( 
        \hat{P} - P(W_i < t) 
    \right)^2 \quad \textbf{where } i \in \{A, B\}
\end{equation}

@@

\begin{equation} \label{eq:payoff_matrices}
    A = 
    \begin{pmatrix}
        U_{1,1}^A & U_{1,2}^A & \dots & U_{1,N_B}^A \\
        U_{2,1}^A & U_{2,2}^A & \dots & U_{2,N_B}^A \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^A & U_{N_A,2}^A & \dots & U_{N_A,N_B}^A \\
    \end{pmatrix}, \,
    B = 
    \begin{pmatrix}
        U_{1,1}^B & U_{1,2}^B & \dots & U_{1,N_B}^B \\
        U_{2,1}^B & U_{2,2}^B & \dots & U_{2,N_B}^B \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^B & U_{N_A,2}^B & \dots & U_{N_A,N_B}^B \\
    \end{pmatrix}
\end{equation}

>>>>

\section{Strategic manipulation of the ED-EMS interface} \label{sec:model_overview}

@@

\begin{itemize}
    \item \( \lambda_2 \): The rate of non-emergency patients that the EMS 
    receives and distributes to EDs
    % TODO: Use consistent notation of non-urgent/non-emergency patients
    \item \( \lambda_{1_i} \): The arrival rate of other patients to ED 
    \(i\in\{A, B\}\)
    \item \( \mu_i \): The service rate of patients at ED \(i\in\{A, B\}\)
    \item \( C_i \): The number of available resources (healthcare 
    professionals) in the ED \(i\in\{A, B\}\)
    % TODO: Make sure that servers are defined the same way throughout
    \item \( T_i \): The action that ED \(i\in\{A, B\} \) chooses to play which 
    corresponds to the threshold at which they do not accept EMS patients.
    \item \( N_i \): The total patient capacity of the ED \(i\in\{A, B\} \)
    \item \( M_i \): The total parking capacity of the ED \(i\in\{A, B\} \)
    \item \( t \): The time target for both EDs
    \item \( \alpha \in [0, 1] \) : Weighted average of blocking time and 
    lost individuals (equation \ref{eq:obj_distributor})
\end{itemize}

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
As defined in section \ref{sec:model_overview}, the interaction between the two
EDs is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). 
This corresponds to reacting to experienced delays.

The queueing systems of the hospitals are designed in such a way where they can
accept two types
of individuals. 
Type 2 individuals are defined as non-urgent patients arriving in ambulances 
that are allocated by the ambulance service and 
type 1 individuals are individuals arriving from other sources 
(e.g walk-in patients, urgent patients from either walk-ins or ambulances).
% TODO: Type 1 - type 2 definition defined on queueing section as well
Each hospital may then choose to block type 2 individuals 
when the hospital reaches a certain capacity. 
The strategy sets for each hospital is the set 
\( \{T \in \mathbb{N} \;|\; 1 \leq T \leq N\} \) where \(N \in\{N_A, N_B\}\) are 
the total capacities of hospitals \(A\) and \(B\). We denote the chosen actions 
from the strategy set as \(T_A, T_B\) and call these \textit{threshold}s.

Both hospitals follow a queueing model with two waiting spaces for 
individuals. 
The first waiting space (i.e. the waiting space of the hospital) is where the 
patients queue right before receiving 
their service and has a queue capacity of \( N - C \), where \(N\) is the total 
capacity of the hospital and \(C\) is the number of healthcare 
professionals able to see them.
% TODO: Should this be healthcare professionals or available beds?
The second waiting space (i.e. the parking space for ambulances) is where 
ambulances, that are sent from the
ambulance service, stay until their patients are allowed to enter the hospital.
The parking space has a capacity of \(M\) and no servers.

@@

% TODO: Maybe only this equation is needed here. Remove the previous two?
\begin{equation}\label{eq:obj_queueing_systems}
    \arg \max_{T_i} \quad -\left( \hat{P} - P(W_i < t) \right)^2
\end{equation}

The third player, the ambulance service, has their own choices to make and 
their own 
goals to satisfy.
The strategy set of the third player is the proportion \(0 \leq p_A \leq 1\) of 
individuals to send to hospital A.
Similarly the proportion of individuals to send to hospital B is given by
\(p_B = 1 - p_A\).
In addition, the ambulance service aims to minimise any potential blockages
that may occur, given the pair of thresholds chosen by the two hospitals.
Thus, its objective is to minimise the blocked time of the individuals 
(\(B_A\) and \(B_B\))
that they send to hospitals \(A\) and \(B\).
Apart from the time being blocked, an additional aspect that may affect the 
decision of the distributor is the proportion of lost individuals 
\(L_A\) and \(L_B\).
Equation \ref{eq:obj_distributor} can be used to capture a mixture 
between the two objectives \(L_i\) and \(B_i\) where \(i \in \{A, B\}\):
\begin{equation}\label{eq:obj_distributor}
    (p_A, p_B) \quad s.t. \quad 
    \alpha L_A(p_A) + (1 - \alpha) B_A(p_A) = 
    \alpha L_B(p_B) + (1 - \alpha) B_B(p_B)
\end{equation}

@@

Hospital \(H_A\) decides on a threshold, then the hospital \(H_B\)
chooses its own threshold, without knowing the strategy of \(H_A\), and finally
the ambulance service makes its choice. 
Note here that the dotted line represents the
fact that \(H_B\) is unaware of the state of the game when making its own 
decisions.

@@

\begin{equation} \label{eq:payoff_entry}
    U_{T_A, T_B}^i = 1 - \left( 
        \hat{P} - P(W_i < t) 
    \right)^2 \quad \textbf{where } i \in \{A, B\}
\end{equation}

@@

\begin{equation} \label{eq:payoff_matrices}
    A = 
    \begin{pmatrix}
        U_{1,1}^A & U_{1,2}^A & \dots & U_{1,N_B}^A \\
        U_{2,1}^A & U_{2,2}^A & \dots & U_{2,N_B}^A \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^A & U_{N_A,2}^A & \dots & U_{N_A,N_B}^A \\
    \end{pmatrix}, \,
    B = 
    \begin{pmatrix}
        U_{1,1}^B & U_{1,2}^B & \dots & U_{1,N_B}^B \\
        U_{2,1}^B & U_{2,2}^B & \dots & U_{2,N_B}^B \\
        \vdots & \vdots & \ddots & \vdots \\
        U_{N_A,1}^B & U_{N_A,2}^B & \dots & U_{N_A,N_B}^B \\
    \end{pmatrix}
\end{equation}
% TODO: Define here the normal form game properly

### 2022-03-09 sections/model_overview/methodology/main.tex
Minor changes
- Minor changes to sections and subsections
- Move figure from game section to queueing section
Restructure game theoretic model section

\section{Methodology}\label{sec:methodology}

The problem defined in section \ref{sec:model_overview} describes a normal-form
game between the decision makers of two hospitals and a third player, the 
ambulance service, that 
decides how to distribute individuals to the systems.
The strategy space of the two hospitals is defined as the possible values
that the threshold parameter can take \(T_i \in [1, N_i]\).
Then, the ambulance service has to decide on the proportion of individuals to 
send to 
each hospital \(p_A \text{ and } p_B\), where \(p_A, p_B \in [0, 1] \)
and \(p_A + p_B = 1\).
In practice this would correspond to a learned behaviour through experience of
waiting at each hospital.
Figure \ref{fig:imperfect_info_game} 
shows a diagrammatic representation of the game to be played and the decisions 
to be made.
As described in Section \ref{sec:model_overview}, hospital \(A\) decides
on a strategy and at the same time, hospital \(B\) chooses its own 
threshold but unaware of the first hospital's choice.
Finally, the ambulance service makes its choice based on the strategies that the 
hospitals chose to play.

@@

The routing matrix holds the values \((p_A, p_B)\) which are the proportion 
of ambulance patients to send to queueing systems \(A\) and \(B\).
Each pair \((p_A, p_B)\) can be calculated using equation 
(\ref{eq:obj_distributor}) and is essentially
a best response to the actions of the hospitals.
Thus, using equation \ref{eq:obj_distributor} for all possible sets of 
thresholds we can get the full routing matrix that consists of the proportions
to send to hospital A (\(p_A\)) and to hospital B (\(p_B\)).

@@

\subsection{Backwards induction}

In order to populate the payoff matrices and the routing matrix the method
of backwards induction is used.
Consider Figure \ref{fig:imperfect_info_game} and the flow of the game that was
described (i.e. \(H_A, H_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow H_A, H_B\)). 
Thus for every pair of strategies \(T_A, T_B\), the proportion of individuals 
that are to be transported to each queueing system is calculated first.

Consider a game where the capacities of the two systems are \(N_A = 4\) and 
\(N_B = 3\).
The strategy space of players \(H_A\) and \(H_B\) would then be 
\(T_A = \{1, 2, 3, 4\}\) and \(T_B = \{1, 2, 3\}\) respectively.
Now, starting from an arbitrary starting point of \(T_A=1\) and \(T_B=1\), the
corresponding entry of the routing matrix the routing matrix can be calculated.
Using equation (\ref{eq:obj_distributor}) for \(T_A=1\) and \(T_B=1\):
\begin{align}
    & \alpha L_A(p_A;T_A=1;T_B=1) + (1 - \alpha) B_A(p_A;T_A=1;T_B=1) = 
    \nonumber \\
    & \alpha L_B(p_B;T_A=1;T_B=1) + (1 - \alpha) B_B(p_B;T_A=1;T_B=1) 
    \label{eq:obj_distributor_1_1}
\end{align}

The values of \(p_A\) and \(p_B\) that satisfy equation 
(\ref{eq:obj_distributor_1_1}) are here found numerically using Brent's 
bisection algorithm~\cite{brent1973algorithms}.
The calculated value of \(p_A\) corresponds to the entry on the first row and 
first column of the routing matrix:

\begin{equation}\label{eq:routing_matrix_1_1}
    R = 
    \begin{pmatrix}
        p_{1,1}^A & - & - \\
        - & - & - \\
        - & - & - \\
        - & - & - \\
    \end{pmatrix}
\end{equation}

In fact all remaining values of the routing matrix can be calculated by
solving \(N \times M\) similar equations using Brent's bisection algorithm.
Therefore, having calculated the entire routing matrix the payoff matrices can
be calculated.
Consider once again the case of \(T_A=1, T_B=1\).

\begin{align}
    U_{1, 1}^A = 1 -\left( 
        \hat{P} - P(W_A(T_A=1, T_B=1) < t) 
    \right)^2 \nonumber
    \\
    U_{1, 1}^B = 1 -\left( 
        \hat{P} - P(W_B(T_A=1, T_B=1) < t) 
    \right)^2 \label{eq:payoff_entry_1_1}
\end{align}

These are the utilities of hospitals \(A\) and \(B\) when both players choose a
strategy of \(T = 1\).
Here \(R\) is the predefined waiting time target to be met by a percentage of 
individuals \( \hat{P} \) and \(P(W_i < t)\) is defined in section
\ref{sec:proportion_within_target}.

\begin{equation}\label{eq:payoff_matrices_1_1}
    A = 
    \begin{pmatrix}
        U_{1, 1}^A & - & - \\
        - & - & - \\
        - & - & - \\
        - & - & - \\
    \end{pmatrix}, \quad
    B = 
    \begin{pmatrix}
        U_{1, 1}^B & - & - \\
        - & - & - \\
        - & - & - \\
        - & - & - \\
    \end{pmatrix}
\end{equation}

Similar to the routing matrix, the above procedure can be repeated for all 
possible values of \(T_A\) and \(T_B\) to populate all entries of the payoff 
matrices.

\subsection{Nash equilibrium}\label{sec:methodology_nash_equilibrium}

Having calculated the payoff matrices \(A\) and \(B\), several algorithms can 
be used to measure some form of emergent behaviour.
One possibility would be to compute the Nash equilibrium
which is the point of the game were both players have no 
incentive to deviate from their played strategies \cite{kreps1989nash}.
In other words their chosen strategies are a best response to each other.

Computation of Nash equilibria can be done in a relatively efficient way using 
the Lemke Howson algorithm~\cite{LemkeHowson}.
Lemke-Howson uses best response polytopes to get one of the Nash equilibrium of
the game. 
Other algorithms exists that will compute all Nash equilibria but 
for large games the computational complexity becomes problematic.
All game theoretic calculations were done in Python using the Nashpy library 
\cite{thenashpyproject}.

\subsection{Learning algorithms}\label{sec:methodology_learning_algorithms}

>>>>

\subsection{Building the game}\label{sec:methodology}

The problem defined in this section describes a normal-form
game between the decision makers of two hospitals and a third player, the 
ambulance service, that 
decides how to distribute individuals to the two systems.
The strategy space of the two hospitals is defined as the possible values
that the threshold parameter can take \(T_i \in [1, N_i]\).
Then, the ambulance service has to decide on the proportion of individuals to 
send to 
each hospital \(p_A \text{ and } p_B\), where \(p_A, p_B \in [0, 1] \)
and \(p_A + p_B = 1\).
In practice this would correspond to a learned behaviour through experience of
waiting at each hospital.
Figure \ref{fig:imperfect_info_game} 
shows a diagrammatic representation of the game to be played and the decisions 
to be made.
As described in section \ref{sec:model_overview}, hospital \(A\) decides
on a strategy and at the same time, hospital \(B\) chooses its own 
threshold but unaware of the first hospital's choice.
Finally, the ambulance service makes its choice based on the strategies that the 
hospitals chose to play.

@@

The routing matrix holds the values \((p_A, p_B)\) which are the proportion 
of ambulance patients to send to queueing systems \(A\) and \(B\).
Each pair \((p_A, p_B)\) can be calculated using equation 
(\ref{eq:obj_distributor}) and is essentially
a best response to the actions of the hospitals.
Thus, using equation (\ref{eq:obj_distributor}) for all possible sets of 
thresholds we can get the full routing matrix that consists of the proportions
to send to hospital A (\(p_A\)) and to hospital B (\(p_B\)).

@@

Consider Figure \ref{fig:imperfect_info_game} and the flow of the game that was
described (i.e. \(H_A, H_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow H_A, H_B\)). 
This is done using backwards induction, by first solving the game from the
ambulance's point of view and then from the hospital's point of view.
Thus, for every pair of strategies \(T_A, T_B\), the values of \(p_A\) and 
\(p_B\) that satisfy equation (\ref{eq:obj_distributor}) are found
numerically using Brent's bisection algorithm~\cite{brent1973algorithms}.
Each pair \((p_A, p_B)\) corresponds to the best response of the ambulance
service to the two hospitals' played strategies.
Finally, using the routing matrix, equation (\ref{eq:payoff_entry}) can also be
used to populate the payoff matrices of the hospitals since we now know the 
arrival rate of each hospital.

Having calculated the payoff matrices \(A\) and \(B\), several algorithms can 
be used to measure some form of the emergent behaviour.
One possibility would be to compute the Nash equilibrium
which is the point of the game were both players have no 
incentive to deviate from their played strategies \cite{kreps1989nash}.
In other words their chosen strategies are a best response to each other.
Computation of Nash equilibria can be done in a relatively efficient way using 
the Lemke Howson algorithm~\cite{LemkeHowson}.
Lemke-Howson uses best response polytopes to get one of the Nash equilibrium of
the game. 
Other algorithms exists that will compute all Nash equilibria but 
for large games the computational complexity becomes problematic.
All game theoretic calculations were done in Python using the Nashpy library 
\cite{thenashpyproject}.

### 2022-03-09 sections/model_overview/numerical_results/main.tex
Restructure game theoretic model section

\subsection{Numerical results}\label{sec:numerical_results}

@@

The most commonly used method for analysing normal form games is the Nash 
equilibrium described in section \ref{sec:methodology_nash_equilibrium}.
Consider the following game:

@@

Nash equilibria is a theoretical measure which can be
inconsistent with intuitive notions about what should be the outcome of a 
game \cite{myerson1978refinements}.
Therefore it might not be the best way to describe human behaviour.
Since the work of Maynard Smith \cite{smith1986evolutionary}, evolutionary game 
theory gives the tools for the emergence of stable behaviour.
One such model that allows for asymmetric payoffs, as is the case above, is 
replicator dynamics described in Section 
(\ref{sec:methodology_learning_algorithms}).
Stable outcomes of this algorithm will correspond to a subset of Nash
equilibria but more importantly, will give a model of emergent behaviour.

@@

Consider a game with two smaller (lower \( N_i, M_i \)) and busier (higher 
\(\lambda_{1_i} \) and \(\lambda_2\)) hospitals with the following set of 
parameters:

>>>>

\subsection{Results}\label{sec:numerical_results}

@@

The most commonly used method for analysing normal form games is the Nash 
equilibrium described in section \ref{sec:methodology}.
Consider the following game:

@@

Nash equilibria is a theoretical measure which can be
inconsistent with intuitive notions about what should be the outcome of a 
game \cite{myerson1978refinements}.
Therefore it might not be the best way to describe human behaviour.
Since the work of Maynard Smith \cite{smith1986evolutionary}, evolutionary game 
theory gives the tools for the emergence of stable behaviour.
One such model that allows for asymmetric payoffs, as is the case above, is 
replicator dynamics described in section 
(\ref{sec:methodology}).
Stable outcomes of this algorithm will correspond to a subset of Nash
equilibria but more importantly, will give a model of emergent behaviour.

@@

% TODO: Make clear that we are speaking about a flooded model here.
Consider a game with two smaller (lower \( N_i, M_i \)) and busier (higher 
\(\lambda_{1_i} \) and \(\lambda_2\)) hospitals with the following set of 
parameters:
